---
title: "R Notebook"
output: html_notebook
---
1.Explore the overall structure of the dataset using “str” and “summary” functions.  

```{r}
data <- read.csv(file="C:/ml/realtor.csv")
```

```{r}
str(data)
```
# 12 Variables and 923159 observations is there.

```{r}
summary(data)
```
2.What is the type of each variable categorical/discrete  or numeric/continuous? For 
categorical/discrete variables, specify whether it is nominal or ordinal. 

categorical-status,street,full_address,city,state,sold_date 
numerical-price,bed,bath,acre_lot,zip_code, house_size 


```{r}
install.packages("tidyverse")
library("tidyverse")
```
3.are there any duplicate observations in the data? If so, remove them. You can use  “duplicated” or 
“unique” functions to answer this question.


```{r}
duplicated(data)
```
#this are duplicated observations.


```{r}
unique(data)
```
```{r}
library(magrittr)
library(dplyr)
data=data %>% distinct()
```

4.Does any of the variables have missing values? Which ones?

```{r}
is.na(data)
```
5.The price variable appears to have some extreme values. Remove the outliers in the “price” variable 
using  the  IQR  method.  IQR  and  quantile  functions  throw  error  if  you  have  NAs  in  your  variable.  Use 
na.rm=TRUE option inside IQR and quantile methods to ignore the missing price values. 

```{r}
IQR <- IQR(data$price,na.rm=TRUE)
```

```{r}
IQR
```
```{r}
summary(data$price)
```
```{r}
left <- 250000-1.5*IQR
left
```

```{r}
right <-800000 +1.5*IQR
right
```
```{r}
x <- data$price
y=x[x<left | x>right]
y
```
```{r}
quantile(data$price,na.rm=TRUE)
```
```{r}
length(y)
```

6.Remove all houses with price less than or equal to 50K
```{r}
data$price[(data$price<=50000), drop = FALSE]
```
```{r}
data$price
```
7.Draw  a  histogram  and  boxplot  of  the  price.  What  can  you  say  about  the  shape  of  price 
variable? Is the price variable positively skewed, symmetric, or negatively skewed? 

```{r}
 boxplot(data$price, main="Boxplot of Prices",
ylab="Price ($)")
```
*Many outliers are expensive for their class.

```{r}
hist(data$price, main = "Histogram of Prices", xlab =
"Price ($)")
```
* The histogram indicates that it is right-skewed.

8.what percentage of the observations are missing for the price variable?

```{r}
mean(is.na(data$price))*100
```

9.Use as.Date method to convert the sold_date variable from character string to date/time type. Then from this date/time object create two new attributes (sold_year) and (sold_month) to store the year and month that the house was sold

```{r}
sold_date_formatted <- as.Date(data$sold_date, format = "%Y-%m-%d")
sold_date_formatted 
```
```{r}
class(sold_date_formatted) 
```
```{r}
sold_date_month <- format(sold_date_formatted, "%m")
sold_date_month 
```
```{r}
data$sold_year<- format(sold_date_formatted, "%Y")
data$sold_year
```
10.convert the “state” attribute to factor then take a summary to see how many observations are there 
for each state. Remove states with only one observation from the data. 

```{r}
data$state<-as.factor(data$state)

data$state
```
```{r}
attach(data)
```

```{r}
summary(data$price)
```
```{r}
library(dplyr)

data<-data %>% group_by(data$state) %>% filter(n()>1)
```
```{r}
summary(data$state)
```

11.Is there a statistically significant difference between the average house price for different states? 
Use appropriate plot and statistic test to answer this question. 


```{r}
plot( data$price,na.rm=TRUE ~ data$state , col="blue")
```

```{r}
oneway.test ( data$price ~ data$state, data=data)

```
12.What is the correlation between house_price and the variables sold_year, house_size, bed, and bath?

```{r}
cor(data$price, data$bath, method = 'spearman',use="pairwise.complete.obs" )
```
```{r}
cor.test(data$price, data$bath, method = 'spearman',use="pairwise.complete.obs" )
```
```{r}
cor(data$price, data$house_size, method = 'spearman',use="pairwise.complete.obs" )
```
```{r}
cor.test(data$price, data$house_size, method = 'spearman',use="pairwise.complete.obs" )
```
```{r}
cor(data$price, data$bed, method = 'spearman',use="pairwise.complete.obs" )
```
```{r}
cor.test(data$price, data$bed, method = 'spearman',use="pairwise.complete.obs" )
```
```{r}
summary(data$sold_year)
```
```{r}
cor(data$price, as.numeric(data$sold_year), method = 'spearman',use="pairwise.complete.obs" )
```
```{r}
cor.test(data$price, as.numeric(data$sold_year), method = 'spearman',use="pairwise.complete.obs" )
```

